---
title: "Qualitative Analisys - BRASFI"
author: "Annie Cavalcante"
format: html
editor: visual
jupyter: python3
editor_options: 
  chunk_output_type: console
---

## Preparing for analysis

```{r, packages}
#|echo: TRUE
#|message: FALSE
library(reticulate)
py_config()

```

```{r, pandas}
#|echo: TRUE
#|message: FALSE
# Install pandas, matplotlib, and wordcloud
py_require(c("pandas", "matplotlib", "wordcloud"))
py_require("nltk")
py_require("openpyxl")


```

## Qualitative analysis

```{python, packages}
#|echo: TRUE
#|message: FALSE
import pandas as pd
import matplotlib.pyplot as plt
from wordcloud import WordCloud
import os
```

The file contains responses to a questionnaire administered via Google Forms. The spreadsheet was reviewed, and multiple-choice and open-ended questions were identified. The main questions identified were:

-   "O que normalmente te impede de participar das atividades da BRASFI?"

-   "O que mais te motivaria a participar ativamente da BRASFI?"

-    "O que podemos ajustar para facilitar sua participação e engajamento?"

-   "Qual tema ou formato de atividade te interessaria mais neste momento?"

-    "Se você pudesse mudar algo na BRASFI com um estalar de dedos, o que seria?"

-    "De que forma você poderia ajudar a potencializar a BRASFI?"

    After identifying the main questions, it was necessary to remove columns that might contain sensitive data.

```{python, file}
#|echo: TRUE
#|message: FALSE

# Load the Excel file
file_path = "/cloud/project/survey_data.xlsx"
df_raw = pd.read_excel(file_path, sheet_name="Respostas ao formulário 1")  

# Remove the last column
df = df_raw.iloc[:, :-1]

# Display the modified DataFrame
print(df)

```

After inspecting the questions and answers provided, we separate them in two categories:

-   Open questions: These are questions that require an answer to be entered.

-   Closed questions: Questions that provide a list of options to choose from, or that require you to select all applicable options.

```{python, closed}
#|echo: TRUE
#|message: FALSE

# Select the 'closed questions' (categorical or multiple choices answers)
colunas_fechadas = [
    "Você atualmente participa de algum setor, núcleo ou frente de trabalho dentro da BRASFI?",
    "Você sente que existe alinhamento entre suas entregas para a BRASFI e o retorno que você esperava (aprendizado, conexões, visibilidade, etc)?",
    "Nos últimos 6 meses, você participou de alguma atividade da BRASFI (Palestras, fóruns, webinars e mentoria)? (Você pode selecionar mais de uma opção)",
    "O formato atual de dias e horários (preferencialmente às quartas, 19h) são acessíveis para sua rotina?",
    "Qual frequência você considera ideal realizar os encontros?",
    "Caso não conseguisse participar da palestra ao vivo, você assistiria à gravação?",
    "Como você avalia os temas discutidos até agora?",
    "Você gostaria de ser ouvido(a) em rodas de conversa ou encontros informais da BRASFI?"
]
```

You can add options to executable code like this

```{python}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
